# Basic Analysis for motorway

First, I will look at one satellite picture to show the general procedure.

According to the Landsat [website](https://landsat.usgs.gov/landsat-surface-reflectance-quality-assessment) the pixel_qa band has a more accurate prediction for identifying clouds, cloud shadows, water, and snow.  
When transforming spatial vector objects to raster objects, I always rely on the fasterize package which is written in C++ and a lot faster than the functions in the raster package.



```{r pixelqatransform,cache=TRUE}
cloud_identification <- 
  raster("E:/landsatscene/LT051940232011072701T1-SC20180926185810/LT05_L1TP_194023_20110727_20180312_01_T1_pixel_qa.tif") 

# crop raster to extent of MV
cloud_identification_cropped <- 
  crop(cloud_identification,
       st_transform(mv,crs=proj4string(cloud_identification)))

# transform MV Polygon to a raster
mv_raster <- 
  fasterize::fasterize(st_transform(mv,
                                    crs=proj4string(cloud_identification)),
                       cloud_identification_cropped)

# set all values that lie outside of MV to NA
cloud_identification_masked <- mask(cloud_identification_cropped,mv_raster)

# extract the values of the cloud raster
cloud_identification_values <- getValues(cloud_identification_masked)

# reference values for not having a clear sight on the ground
none_clear_values <- c(96, 112,160,176,224,68,72,80,96, 
                       112,132,136,144,160,176,224,72,
                       136,68,132,80,112,144,176)

# Create a mask vector. NA for each pixel outside of MV, 0 for if its inside
# but not a clear view, and 1 if its inside MV and there is a clear view
cloud_identification_values_2 <- 
  ifelse(cloud_identification_values %in% none_clear_values,0,
         ifelse(is.na(cloud_identification_values),NA,1))

# create raster from mask vector
cloud_identification_2 <- setValues(cloud_identification_masked,
                                    cloud_identification_values_2) 

# transform motorways to a SpatialPolygon by buffering 1 meter around the lines
motorways_sf <- spTransform(motorways,
                            CRS(proj4string(cloud_identification_2))) %>% 
  gBuffer(width = 2) %>% 
  st_as_sf()

# transform roads to a raster
motorways_raster <- fasterize::fasterize(motorways_sf,
                     cloud_identification_2)

covered_by_road <- mask(cloud_identification_2,motorways_raster) %>% 
  getValues() %>% 
  is.na() %>% 
  `!` %>% 
  which()

cloud_identification_values_3 <- ifelse(is.na(cloud_identification_values_2),0,cloud_identification_values_2)
```

```{r tableofcloud,echo=FALSE,cache=FALSE}
Rcpp::sourceCpp("cppsource/table_cpp.cpp")
cbind(
  table_cpp(cloud_identification_values_2[!is.na(cloud_identification_values_2)]) %>% 
  do.call(rbind,.),c(sum(is.na(cloud_identification_values_2)),NA)) %>% 
  t %>% 
  as.data.frame %>% 
  mutate(lengths = round(100*lengths/sum(lengths),2)) %>% 
  rename(Percentage = lengths,
         Values = values) %>% 
  kable(caption="Coverage for different values. `NA` are the rims of the picture, `0` is cloud covered, and `1` is free sight.")
```

The following code chunk loads each band, crops it to the extent of MV, and sets all pixel to NA that either have no undisturbed sight or lie outside of MV.
The values are bound together in a rectengular data.frame with the added information whether the specific pixel is a motorway or not. All NAs are subsequently removed from the data.frame and all values are mean normalized.

```{r rasterbrick1,cache=TRUE}
listed_raster <- list.files("E:/landsatscene/LT051940232011072701T1-SC20180926185810/",pattern="band[0-9].tif$",full.names = TRUE)

values_df <- lapply(listed_raster,function(ra1_path) {
  
  ra1 <- raster(ra1_path)
  
  ra1_cropped <- crop(ra1,st_transform(mv,crs=proj4string(ra1)))
  
  ra1_cropped_masked <- mask(ra1_cropped,mv_raster)
  
  ra1_cropped_masked_cloudless <- setValues(ra1_cropped_masked,ifelse(getValues(ra1_cropped_masked)*cloud_identification_values_3==0,NA,getValues(ra1_cropped_masked)))
  
  getValues(ra1_cropped_masked_cloudless)
  
}) %>% 
  do.call(cbind,.) %>% 
  as.data.frame()

values_df$motorway <- 0
values_df$motorway[covered_by_road] <- 1

identify_na <- sapply(paste0("V",1:6),function(x) {
   values_df[,x] %>% 
  is.na() %>% 
  which()
}) %>%
  unlist() %>% 
  unique()

if (length(identify_na) > 0) {
  values_df <- values_df %>% 
    dplyr::slice(-identify_na)
}

norm_mean <- function(x) {
  (x-mean(x))/(max(x)-min(x))
}

values_df[,paste0("V",1:6)] <- values_df[,paste0("V",1:6)] %>% 
  apply(2,norm_mean)

```



The `train_and_test_data` function is creating a train and test set. If `replace_motorway` is set to `TRUE` the same amount of motorway pixel will be sampled for the training data as of no motorway pixel. This way, the distribution will 

```{r functions1}
train_and_test_data <- function(seed,df,trainsize=0.8,
                                replace_motorway=TRUE,
                                sample_size=1e4,
                                sample_test=1e4,
                                factorize_dependent=FALSE) {
  set.seed(seed)
  if (factorize_dependent) df$motorway <- as.factor(df$motorway)
  motorway_pixel <- which(df$motorway==1)
  no_motorway_pixel <- which(df$motorway==0)
  
  sampled_motorway_pixel <- sample(motorway_pixel,
                                   round(trainsize*length(motorway_pixel)))
  sampled_no_motorway_pixel <- sample(no_motorway_pixel,
                                      round(trainsize*length(no_motorway_pixel)))
  
  test_motorway_pixel <- motorway_pixel[!motorway_pixel %in% 
                                          sampled_motorway_pixel]
  test_no_motorway_pixel <- no_motorway_pixel[!no_motorway_pixel %in% 
                                                sampled_no_motorway_pixel]
  
  values_df_train <- df[c(sampled_motorway_pixel,sampled_no_motorway_pixel),]
  values_df_test <- df[c(test_motorway_pixel,test_no_motorway_pixel),]
  
  
  
  # train
  if (replace_motorway) {
    selected_motorway <- sample(sampled_motorway_pixel,sample_size,replace=TRUE)
    selected_non_motorway <- sample(sampled_no_motorway_pixel,sample_size)
  } else {
    selected_motorway <- sampled_motorway_pixel
    selected_non_motorway <- sample(sampled_no_motorway_pixel,sample_size)
  }
  
  subset <- c(selected_non_motorway,selected_motorway)
  values_df_model <- dplyr::slice(df,subset)
  
  # test
  subset_test <- c(test_motorway_pixel,sample(test_no_motorway_pixel,
                                              sample_test))
  values_df_test <- dplyr::slice(df,subset_test)
  
  
  list(train=values_df_model,
       test=values_df_test)
}
predicted <- function(data) {
    data %>% 
    dplyr::rename(Type=motorway) %>% 
    group_by(Type) %>%
    summarise(Value = sum(equal)/n()) %>% 
    mutate(Type=ifelse(Type==1,"Motorway","None Motorway"))
}

```


```{r randomnumbers}
set.seed(15)
random_number <- sample(1e5,100)
```


## Ordinary Least Squares

```{r linearmodel,echo=TRUE,eval=FALSE}
generated <- lapply(random_number,function(x) {
  
  train_test_list <- train_and_test_data(seed = x,
                                         df = values_df,
                                         trainsize = 0.8,
                                         sample_size = 1e4,
                                         sample_test = 1e4)
  values_df_model <- train_test_list$train
  values_df_test <- train_test_list$test
  
  # model
  model_linear <- lm(motorway~V1+V2+V3+V4+V5+V6,data=values_df_model) 
  
  # predict
  values_df_test$predicted <- predict(model_linear,values_df_test)
  
  values_df_test <- values_df_test %>% 
    mutate(predicted_binary = as.numeric(predicted>0.5),
           equal = as.numeric(motorway==predicted_binary))
  
  
  values_df_test %>% 
    predicted()
    
}) %>% 
  do.call(rbind,.)
```

```{r saveols,eval=FALSE,echo=FALSE}
write.csv(generated,"output/one_scene_model_eval/ols.csv",row.names = FALSE)

```

## Random Forest

```{r randomforest,eval=FALSE,echo=TRUE}
generated_rf <- lapply(random_number,function(x) {
  values_df$motorway <- as.factor(values_df$motorway)
  train_test_list <- train_and_test_data(seed = x,
                                         df = values_df,
                                         trainsize = 0.8,
                                         replace_motorway = FALSE,
                                         sample_size = 1e3,
                                         sample_test = 1e4)
  values_df_model <- train_test_list$train
  values_df_test <- train_test_list$test
  
  # train
  model_rf<-randomForest(
    formula = motorway~V1+V2+V3+V4+V5+V6,
    data= values_df_model,
    ntree=500,
    importance=TRUE)#,

  # predict
  values_df_test$predicted <- predict(model_rf,values_df_test %>% 
                                        dplyr::select(paste0("V",1:6),motorway))
  
  values_df_test2 <- values_df_test %>% 
    mutate(equal = as.numeric(motorway==predicted))
  
  values_df_test2 %>% 
    predicted()
}) %>% 
  do.call(rbind,.)


```

```{r saverf,eval=FALSE,echo=FALSE}
write.csv(generated_rf,"output/one_scene_model_eval/randomforest.csv",row.names = FALSE)

```

## XGBoost

```{r xgboost,eval=FALSE}
generated_xgb <- lapply(random_number,function(x) {
  
  train_test_list <- train_and_test_data(seed = x,
                                         df = values_df,
                                         trainsize = 0.8,
                                         replace_motorway = FALSE,
                                         sample_size = 1e3,
                                         sample_test = 1e4)
  values_df_model <- train_test_list$train
  values_df_test <- train_test_list$test
  
  train_dat<-xgb.DMatrix(
    data = values_df_model %>% 
      dplyr::select(paste0("V",1:6)) %>% 
      as.matrix(),
    label = values_df_model %>% 
      dplyr::select(motorway) %>% 
      unlist)
  
  
  model_xgboost<-xgboost(data = train_dat, # the data   
                         nrounds=1000,
                         eta=0.025,
                         subsample=0.63,
                         lamba=0.001,
                         objective = "binary:logistic")  
  
  test_dat<-xgb.DMatrix(
    data = values_df_test %>% 
      dplyr::select(paste0("V",1:6)) %>% 
      as.matrix(),
    label = values_df_test %>% 
      dplyr::select(motorway) %>% 
      unlist)
  
  values_df_test$predicted <- predict(model_xgboost,test_dat)
  
  values_df_test2 <- values_df_test %>% 
    mutate(predicted_binary = as.numeric(predicted>0.5),
           equal = as.numeric(motorway==predicted_binary))
  
  
  values_df_test2 %>% 
    predicted()
}) %>% 
  do.call(rbind,.)
```

```{r savexgb,eval=FALSE,echo=FALSE}
write.csv(generated_xgb,"output/one_scene_model_eval/xgboost.csv",row.names = FALSE)

```

## Neural Network

```{r keras,eval=FALSE}
generated_keras <- lapply(random_number,function(x) {
  
  train_test_list <- train_and_test_data(seed = x,
                                         df = values_df,
                                         trainsize = 0.8,
                                         replace_motorway = TRUE,
                                         sample_size = 2e4,
                                         sample_test = 1e4)
  
  values_df_model <- train_test_list$train
  values_df_test <- train_test_list$test
  
  model <- keras_model_sequential()  
  model %>% 
    layer_dense(units = 256, activation = 'elu',
                input_shape = c(6),kernel_regularizer = regularizer_l2(0.001)) %>%
    layer_dense(units = 128, activation = 'elu',
                input_shape = c(6),kernel_regularizer = regularizer_l2(0.001)) %>%
    
    layer_dense(units = 64, activation = 'elu',
                input_shape = c(6),kernel_regularizer = regularizer_l2(0.0001)) %>%
    layer_dense(units = 32, activation = 'elu',
                input_shape = c(6),kernel_regularizer = regularizer_l2(0.0001)) %>%
    layer_dense(units = 16,activation = 'elu',
                input_shape = c(6),kernel_regularizer = regularizer_l2(0.0001)) %>%
    layer_dense(units = 8, activation = 'elu',
                kernel_regularizer = regularizer_l2(0.0001)) %>% 
    layer_dense(units = 4, activation = 'elu',
                kernel_regularizer = regularizer_l2(0.0001)) %>% 
    layer_dense(units = 3, activation = 'elu',
                kernel_regularizer = regularizer_l2(0.0001)) %>% 
    layer_dense(units = 2, activation = 'sigmoid')
  
  model %>% 
    compile(
      loss = 'binary_crossentropy',
      optimizer = 'adam',
      metrics = c('accuracy'))
  
  k_set_value(model$optimizer$lr, 0.00001)
  k_set_value(model$optimizer$decay, 0.00001/100)
  
  
  history <- model %>%
    fit(
      x = values_df_model %>% 
        dplyr::select(paste0("V",1:6)) %>% 
        apply(2,as.numeric),
      y = values_df_model %>% 
        dplyr::select(motorway) %>% 
        apply(2,as.numeric)%>% 
        to_categorical(),
      epochs = 60, batch_size = 30, 
      validation_split = 0.1)
  
  predicted <- model %>% 
    predict(values_df_test %>% 
              dplyr::select(paste0("V",1:6)) %>%
              apply(2,as.numeric)) 
  values_df_test$predicted <- predicted[,2]  
  
  
  values_df_test2 <- values_df_test %>% 
    mutate(predicted_binary = as.numeric(predicted>0.5),
           equal = as.numeric(motorway==predicted_binary))
  
  
  values_df_test2 %>% 
    predicted()
}) %>% 
  do.call(rbind,.)
```


```{r savekeras,eval=FALSE,echo=FALSE}
write.csv(generated_keras,"output/one_scene_model_eval/keras.csv",row.names = FALSE)

```

## Convolutional Neural Network

```{r prepforcnn,eval=FALSE}

listed_raster <- list.files("E:/landsatscene/LT051940232011072701T1-SC20180926185810/",pattern="band[0-9].tif$",full.names = TRUE)

norm_mean <- function(x) {
  (x-mean(x,na.rm=TRUE))/(max(x,na.rm=TRUE)-min(x,na.rm=TRUE))
}

ra1_path <- listed_raster[1]

brick <- lapply(listed_raster,function(ra1_path) {
  # ra1 <- raster(ra1_path)
  ve1 <- velox(ra1_path)
  
  ve1$crop(mv_st %>% 
             extent %>% 
             as.vector())
  
  ra1_cropped <- ve1$as.RasterLayer()
  
  ra1_cropped_masked <- mask(ra1_cropped,mv_raster)
  
  ra1_cropped_masked_values <- getValues(ra1_cropped_masked)
  
  cloudless_values <- ra1_cropped_masked_values *cloud_identification_values_3
  
  ra1_cropped_masked_cloudless_values <-
    ifelse_cpp(cloudless_values,0,0,NA,ra1_cropped_masked_values)
    
  
  ra1_cropped_masked_cloudless_values_norm <- norm_mean(ra1_cropped_masked_cloudless_values)
  
  setValues(ra1_cropped_masked,ra1_cropped_masked_cloudless_values_norm)
    
}) %>% 
  brick

# aggregate
# to polygon
# extract from each band
# return as array in list

# aggregate raster to get mask to crop -- faster with velox
b1_velox <- velox(brick[[1]])
b1_agg <- b1_velox$copy()
b1_agg$aggregate(factor=c(5,5),aggtype="sum")
agg_raster <- b1_agg$as.RasterLayer()
agg_raster %>% getValues() %>% is.na %>% which %>% length
agg_raster %>% getValues() %>% length
# transform mask to spatialpolygon

b1_agg_poly <- agg_raster %>% 
  as("SpatialPolygons")

# again, do it in velox because its way faster
brick_velox <- velox(brick)


array_of_images <- laply(1:length(b1_agg_poly),function(y) {
  if (y%%1000==0) message(y)
  b1 <- brick_velox$copy()
  b1$crop(b1_agg_poly[y,])
  laply(1:6,function(x) b1$as.matrix(x))
})

# remove with NA
del <- lapply(1:dim(array_of_images)[1],function(x) array_of_images[x,,,] %>% 
                is.na %>% 
                any) %>% 
  unlist %>%
  which
array_of_images <- array_of_images[-del,,,]
# for Yd
motorways_raster01 <- setValues(motorways_raster,ifelse(is.na(getValues(motorways_raster)),0,1))
motorways_velox <- velox(motorways_raster01)
motorways_velox$aggregate(factor=c(5,5),aggtype="max")

Y <- motorways_velox$extract(b1_agg_poly,fun=max)

Y <- Y[-del,]

Y_cat <- to_categorical(Y %>% as.vector())
```


```{r cnn1,eval=FALSE}
pos <- which(Y==1)

generated_cnn <- lapply(random_number,function(x) {
  
  set.seed(x)
  pos_train <- sample(pos,round(length(pos)*0.8)) %>% rep(8)
  pos_test <- pos[!pos%in%pos_train] %>% rep(8)
  
  neg <- which(Y==0)
  neg_train <- sample(neg,length(pos_train))
  neg_test <- sample(neg[!neg%in%neg_train],length(pos_test))
  
    y_train <- Y_cat[c(pos_train,neg_train) %>% 
                     sort,]
  # y_val <- Y_cat[c(pos_val,neg_val) %>% 
                   # sort,]
  y_test <- Y_cat[c(pos_test,neg_test) %>% 
                    sort,]
  x_train <- array_of_images[c(pos_train,neg_train) %>% 
                               sort,,,]
  # x_val <- array_of_images[c(pos_val,neg_val) %>% 
                             # sort,,,]
  x_test <- array_of_images[c(pos_test,neg_test) %>% 
                              sort,,,]  
  
  model<-keras_model_sequential()
  
  model %>%  
    layer_conv_2d(filters=5,kernel_size=5,data_format="channels_first", 
                  padding="same",input_shape=c(6,5,5),activation = "elu"
                  ,kernel_regularizer = regularizer_l2(0.01)
    ) %>%  
    layer_conv_2d(filter=10,kernel_size=5,activation="elu",padding="same",
                  kernel_regularizer = regularizer_l2(0.0001))  %>%
    layer_max_pooling_2d(pool_size=c(3,3)) %>%  
    layer_flatten() %>%  
    layer_dense(units = 64,activation = "elu"
                ,kernel_regularizer = regularizer_l2(0.0001)
    ) %>%  
    layer_dense(units = 32,activation = "elu"
                ,kernel_regularizer = regularizer_l2(0.0001)
    ) %>%  
    layer_dense(units = 16,activation = "elu"
                ,kernel_regularizer = regularizer_l2(0.0001)
    ) %>%  
    layer_dense(units = 8,activation = "elu"
                ,kernel_regularizer = regularizer_l2(0.0001)
    ) %>%  
    
    layer_dense(units = 4,activation = "elu"
                ,kernel_regularizer = regularizer_l2(0.0001)
    ) %>%  
    layer_dense(units = 3,activation = "elu"
                ,kernel_regularizer = regularizer_l2(0.0001)
    ) %>%  
    layer_dense(2,activation = "sigmoid") 
  
  
  model %>%
    compile(loss="categorical_crossentropy",
            optimizer=optimizer_rmsprop(lr = 0.0001,decay = 1e-8),
            # optimizer_adam( lr= 0.0001 , decay = 1e-8 ),
            metrics = "accuracy")
  
  
  
  model %>%
    fit( x_train,y_train ,batch_size=32,
                 epochs=60,#validation_data = list(x_val, y_val),
                 shuffle=TRUE)
  
  
  predicted <- model %>% 
    predict(x_test) 
  
  
  pred <- data.frame(orig=y_test[,2],pred=ifelse(predicted[,2]>0.5,1,0)) %>% 
    mutate(res=apply(.,1,function(x) {
      if (x[1]==x[2]) return(1)
      0
    })) %>% 
    group_by(orig,res) %>% 
    dplyr::summarize(n=dplyr::n())
  
  data.frame(Type=c("none_motorway_accuracy","motorway_accuracy"),
             Value = c(pred$n[2]/sum(pred$n[1:2]),
                       pred$n[4]/sum(pred$n[3:4])))
  
  
}) %>% 
  do.call(rbind,.)
```

```{r cnn_save,eval=FALSE,echo=FALSE}
write.csv(generated_cnn,"output/one_scene_model_eval/cnn.csv",row.names = FALSE)

```





## Evaluation
```{r mergesims,eval=FALSE}
all_simulations <- list(OLS=generated,RandomForest=generated_rf,XGBoost=generated_xgb,Keras=generated_keras,CNN=generated_cnn)
```

```{r reloadsims,echo=FALSE}
methods <- c("OLS","RandomForest","XGBoost","Keras","CNN")
all_simulations <- lapply(methods,function(x) {
  list.files("output/one_scene_model_eval/",full.names=TRUE,pattern=x %>% 
               tolower) %>% 
     read.csv() %>% 
      mutate(Type= ifelse(Type=="motorway_accuracy","Motorway","None Motorway")) 
})



names(all_simulations) <- methods
```

Before evaluating, I remove one prediction of the neural network that predicts that all pixels are none road pixels.
```{r allsimulationsdf}
all_simulations_df <- names(all_simulations) %>% 
  lapply(function(name) {
    dat <- all_simulations[[name]]
    dat$Method <- name
    dat %>% 
      filter(!is.na(Value))
  }) %>% 
  do.call(rbind,.) %>% 
  filter(Value>0&Value<1)

nrow(all_simulations_df)/2
```


Figure \@ref(fig:evalmodelsfor1scene) shows the accuracy for four different methods. All approaches yield similar results with the average accuracy for motorway prediction lying between 68% and 90%. The accuracy of predicting the none motorway pixels is slightly higher and the difference between the methods is more obvious.  
In general OLS performs less well than the other three and the neural network has the greates span and even has some predictions below 70% for motorway. Random Forest and XGBoost yield the best results, especially when it comes to the "none motorway" prediction were both are consistently above 90%.


```{r evalmodelsfor1scene,fig.align="center",fig.cap="Comparison between OLS, Random Forest, XGBoost, and neural network (Keras)."}
all_simulations_df %>% 
  group_by(Method,Type) %>% 
  summarise(mean=round(mean(Value),4)*100) %>% 
spread(Type,mean) %>% 
     kable(caption = "Accuracy for motorway and none motorway pixels for different methods in percent.")

all_simulations_df %>% 
  ggplot() +
  geom_violin(aes(x=Type,y=Value,fill=as_factor(Method))) +
  ylab("Accuracy") +
  guides(fill=guide_legend(title="Methods"))

```

