# Preparation for a basic analysis

First, I will look at one satellite picture to show the general procedure. 


According to the Landsat [website](https://landsat.usgs.gov/landsat-surface-reflectance-quality-assessment) the pixel_qa band has a more accurate prediction for identifying clouds, cloud shadows, water, and snow.  
When transforming spatial vector objects to raster objects, I always rely on the fasterize package which is written in C++ and a lot faster than the functions in the raster package.



```{r pixelqatransform,cache=TRUE}
cloud_identification <- 
  raster("E:/landsatscene/LT051940232011072701T1-SC20180926185810/LT05_L1TP_194023_20110727_20180312_01_T1_pixel_qa.tif") 

# crop raster to extent of MV
cloud_identification_cropped <- 
  crop(cloud_identification,
       st_transform(mv,crs=proj4string(cloud_identification)))

# transform MV Polygon to a raster
mv_raster <- 
  fasterize::fasterize(st_transform(mv,
                                    crs=proj4string(cloud_identification)),
                       cloud_identification_cropped)

# set all values that lie outside of MV to NA
cloud_identification_masked <- mask(cloud_identification_cropped,mv_raster)

# extract the values of the cloud raster
cloud_identification_values <- getValues(cloud_identification_masked)

# reference values for not having a clear sight on the ground
none_clear_values <- c(96, 112,160,176,224,68,72,80,96, 
                       112,132,136,144,160,176,224,72,
                       136,68,132,80,112,144,176)

# Create a mask vector. NA for each pixel outside of MV, 0 for if its inside
# but not a clear view, and 1 if its inside MV and there is a clear view
cloud_identification_values_2 <- 
  ifelse(cloud_identification_values %in% none_clear_values,0,
         ifelse(is.na(cloud_identification_values),NA,1))

# create raster from mask vector
cloud_identification_2 <- setValues(cloud_identification_masked,
                                    cloud_identification_values_2) 

# transform motorways to a SpatialPolygon by buffering 1 meter around the lines
motorways_sf <- spTransform(motorways,
                            CRS(proj4string(cloud_identification_2))) %>% 
  gBuffer(width = 2) %>% 
  st_as_sf()

# transform roads to a raster
motorways_raster <- fasterize::fasterize(motorways_sf,
                     cloud_identification_2)

covered_by_road <- mask(cloud_identification_2,motorways_raster) %>% 
  getValues() %>% 
  is.na() %>% 
  `!` %>% 
  which()

cloud_identification_values_3 <- ifelse(is.na(cloud_identification_values_2),0,cloud_identification_values_2)
```

```{r tableofcloud,echo=FALSE,cache=FALSE}
Rcpp::sourceCpp("cppsource/table_cpp.cpp")
cbind(
  table_cpp(cloud_identification_values_2[!is.na(cloud_identification_values_2)]) %>% 
  do.call(rbind,.),
  c(sum(is.na(cloud_identification_values_2)),NA)) %>% 
  t %>% 
  as.data.frame() %>% 
  mutate(lengths = round(100*lengths/sum(lengths),2)) %>% 
  dplyr::rename(Percentage = lengths,
         Values = values) %>% 
  kable(caption="Coverage for different values. `NA` are the rims of the picture, `0` is cloud covered, and `1` is free sight.")
```

The following code chunk loads each band, crops it to the extent of MV, and sets all pixel to NA that either have no undisturbed sight or lie outside of MV.
The values are bound together in a rectengular data.frame with the added information whether the specific pixel is a motorway or not. All NAs are subsequently removed from the data.frame and all values are mean normalized.

```{r rasterbrick1,cache=TRUE}
listed_raster <- list.files("E:/landsatscene/LT051940232011072701T1-SC20180926185810/",pattern="band[0-9].tif$",full.names = TRUE)

values_df <- lapply(listed_raster,function(ra1_path) {
  
  ra1 <- raster(ra1_path)
  
  ra1_cropped <- crop(ra1,st_transform(mv,crs=proj4string(ra1)))
  
  ra1_cropped_masked <- mask(ra1_cropped,mv_raster)
  
  ra1_cropped_masked_cloudless <- setValues(ra1_cropped_masked,ifelse(getValues(ra1_cropped_masked)*cloud_identification_values_3==0,NA,getValues(ra1_cropped_masked)))
  
  getValues(ra1_cropped_masked_cloudless)
  
}) %>% 
  do.call(cbind,.) %>% 
  as.data.frame()

values_df$motorway <- 0
values_df$motorway[covered_by_road] <- 1

identify_na <- sapply(paste0("V",1:6),function(x) {
   values_df[,x] %>% 
  is.na() %>% 
  which()
}) %>%
  unlist() %>% 
  unique()

if (length(identify_na) > 0) {
  values_df <- values_df %>% 
    dplyr::slice(-identify_na)
}

norm_mean <- function(x) {
  (x-mean(x))/(max(x)-min(x))
}

values_df[,paste0("V",1:6)] <- values_df[,paste0("V",1:6)] %>% 
  apply(2,norm_mean)

```




