# Evaluation of Basic Models
```{r mergesims,eval=FALSE}
all_simulations <- list(OLS=generated,logit=generated_glm,
                      RandomForest=generated_rf,XGBoost=generated_xgb,
                      Keras=generated_keras,CNN=generated_cnn)
```

```{r reloadsims,echo=FALSE}
methods <- c("OLS","logit","RandomForest","XGBoost","Keras","CNN")
all_simulations <- lapply(methods,function(x) {
  list.files("output/one_scene_model_eval/",full.names=TRUE,pattern=x %>% 
               tolower) %>% 
     read.csv() %>% 
      mutate(Type= ifelse(Type=="motorway_accuracy","Motorway","None Motorway")) 
})



names(all_simulations) <- methods
```

Before evaluating, I remove one prediction of the neural network that predicts that all pixels are none road pixels.
```{r allsimulationsdf}
all_simulations_df <- names(all_simulations) %>% 
  lapply(function(name) {
    dat <- all_simulations[[name]]
    dat$Method <- name
    dat %>% 
      filter(!is.na(Value))
  }) %>% 
  do.call(rbind,.) %>% 
  filter(Value>0&Value<1)

nrow(all_simulations_df)/2
```


Figure \@ref(fig:evalmodelsfor1scene) shows the accuracy for six different methods. Most approaches yield similar results with the average accuracy for motorway prediction lying between 68% and 90%. The CNN has a high variety of results which might be due to the fact that the size of the data is a lot smaller than it is for the other 5 methods.  
The accuracy of predicting the none motorway pixels is slightly higher and the difference between the methods is more obvious.  
Surprisingly, the logit is outperformed by the other non-CNN methods for the prediction of motorways. Also, the OLS performs less well than the neural network, XGBoost, and Random Forest. The neural network has the second highest variability and even has some predictions below 70% for motorways. Random Forest and XGBoost yield the best results, especially when it comes to the "none motorway" prediction were both are consistently above 90%.


```{r evalmodelsfor1scene,fig.align="center",fig.cap="Comparison between OLS, Random Forest, XGBoost, and neural network (Keras)."}
all_simulations_df %>% 
  group_by(Method,Type) %>% 
  dplyr::summarise(mean=round(mean(Value),4)*100) %>% 
  spread(Type,mean) %>% 
  kable(caption = "Accuracy for motorway and none motorway pixels for different methods in percent.")

all_simulations_df %>% 
  ggplot() +
  geom_violin(aes(x=Type,y=Value,fill=as_factor(Method))) +
  ylab("Accuracy") +
  guides(fill=guide_legend(title="Methods"))

```
